Properties and Bayes fitting of restricted Boltzmann machines

### 2017-06-13 writing/draft.Rmd
small tweeks

```{r libraries_options, echo=FALSE, message=FALSE, warnings=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

@@

```{r param-table}
old_pen <- options()$scipen
options("scipen" = 100)
sample.params %>% 
  ungroup %>% 
  filter(!near_hull) %>% select(starts_with("v"), starts_with("h"), starts_with("theta"), -H, -V) %>%
  rename(`$\\theta_{v1}$` = v1,
         `$\\theta_{v2}$` = v2,
         `$\\theta_{v3}$` = v3,
         `$\\theta_{v4}$` = v4,
         `$\\theta_{h1}$` = h1,
         `$\\theta_{h2}$` = h2,
         `$\\theta_{h3}$` = h3,
         `$\\theta_{h4}$` = h4,
         `$\\theta_{11}$` = theta11,
         `$\\theta_{12}$` = theta12,
         `$\\theta_{13}$` = theta13,
         `$\\theta_{14}$` = theta14,
         `$\\theta_{21}$` = theta21,
         `$\\theta_{22}$` = theta22,
         `$\\theta_{23}$` = theta23,
         `$\\theta_{24}$` = theta24,
         `$\\theta_{31}$` = theta31,
         `$\\theta_{32}$` = theta32,
         `$\\theta_{33}$` = theta33,
         `$\\theta_{34}$` = theta34,
         `$\\theta_{41}$` = theta41,
         `$\\theta_{42}$` = theta42,
         `$\\theta_{43}$` = theta43,
         `$\\theta_{44}$` = theta44) %>%
  gather("Parameter", "Value") %>%
  mutate(Value = paste0("$", round(Value, 7), "$")) -> tbl
options("scipen" = old_pen)

cbind(tbl[1:8,], tbl[9:16,], tbl[17:24,]) %>%
  kable(caption = "\\label{tab:theta}Parameters used to fit a test case with $V = H = 4$. This parameter vector was chosen as a sampled value of $\\boldsymbol \\theta$ that was not near the convex hull of the sufficient statistics for a grid point in Figure \\ref{fig:degen_plots} with $< 5$\\% near-degeneracy.", align = rep(c("r", "r"), times = 3))
```

@@

\begin{table}[ht]
\centering
\caption{The values used for the hyperparameters for all three fitting methods. A rule of thumb is imposed which decreases prior variances for the model parameters as the size of the model increases and also shrinks $\boldsymbol \theta_{interaction}$ more than $\boldsymbol \theta_{main}$. The common $C$ defining $C_1$ and $C_2$  in the BwTPLV method is chosen by tuning.}
\label{tab:hyperparam}
\begin{tabular}{|l|c|c|}
\hline 
Method & Hyperparameter & Value \\ 
\hline \hline
\multirow{2}{*}{BwTPLV} & $C_1$ & $\frac{C}{n}\frac{1}{\nh + \nv}$ \\
 & $C_2$ & $\frac{C}{n}\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNLV} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNML} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\end{tabular}
\end{table}

@@

```{r m_eff_table}
M_eff_table %>%
  gather(Image, M_eff, -Model) %>%
  spread(Model, M_eff) %>%
  mutate(Outcome = as.numeric(Image)) %>%
  select(Outcome, BwTNLV, BwTNML) %>%
  arrange(Outcome) -> M_eff_tmp

cbind(M_eff_tmp[1:8,], M_eff_tmp[9:16,]) %>%
  kable(digits = 2, caption = paste0("\\label{tab:m_eff}The effective sample sizes for a chain of length $M = 1000$ regarding all $16$ probabilities for possible vector outcomes of visibles. BwTNLV would require at least $", round(min(M_eff_tmp[, "BwTNML"])/min(M_eff_tmp[, "BwTNLV"]), 1), "$ times as many MCMC iterations to achieve the same amount of effective information about the posterior distribution."))
```

The effective sample sizes for a chain of length $M = 1000$ for inference about each of the $2^4 = 16$ model probabilities are presented in Table \ref{tab:m_eff}. These range from  $`r round(min(M_eff_tmp[, "BwTNML"]), 2)`$ to $`r round(max(M_eff_tmp[, "BwTNML"]), 2)`$ for BwTNML, while BwTNLV only yields between $`r round(min(M_eff_tmp[, "BwTNLV"]), 2)`$ and $`r round(max(M_eff_tmp[, "BwTNLV"]), 2)`$ effective draws. Thus, BwTNLV would require at least $`r round(min(M_eff_tmp[, "BwTNML"])/min(M_eff_tmp[, "BwTNLV"]), 1)`$ times as many iterations to be run of the MCMC chain in order to achieve the same amount of effective information about the posterior distribution. For this reason, consistent with the ACF results in Figure \ref{fig:acf}, BwTNLV does not seem to be an effective method for fitting the RBM if computing resources are at all limited.

>>>>

```{r libraries_options, echo=FALSE, message=FALSE, warnings=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(xtable)

@@

```{r param-table, results='asis'}
old_pen <- options()$scipen
options("scipen" = 100)
sample.params %>% 
  ungroup %>% 
  filter(!near_hull) %>% select(starts_with("v"), starts_with("h"), starts_with("theta"), -H, -V) %>%
  rename(`$\\theta_{v1}$` = v1,
         `$\\theta_{v2}$` = v2,
         `$\\theta_{v3}$` = v3,
         `$\\theta_{v4}$` = v4,
         `$\\theta_{h1}$` = h1,
         `$\\theta_{h2}$` = h2,
         `$\\theta_{h3}$` = h3,
         `$\\theta_{h4}$` = h4,
         `$\\theta_{11}$` = theta11,
         `$\\theta_{12}$` = theta12,
         `$\\theta_{13}$` = theta13,
         `$\\theta_{14}$` = theta14,
         `$\\theta_{21}$` = theta21,
         `$\\theta_{22}$` = theta22,
         `$\\theta_{23}$` = theta23,
         `$\\theta_{24}$` = theta24,
         `$\\theta_{31}$` = theta31,
         `$\\theta_{32}$` = theta32,
         `$\\theta_{33}$` = theta33,
         `$\\theta_{34}$` = theta34,
         `$\\theta_{41}$` = theta41,
         `$\\theta_{42}$` = theta42,
         `$\\theta_{43}$` = theta43,
         `$\\theta_{44}$` = theta44) %>%
  gather("Parameter", "Value") %>%
  mutate(Value = paste0("$", round(Value, 7), "$")) -> tbl
options("scipen" = old_pen)

cbind(tbl[1:8,], tbl[9:16,], tbl[17:24,]) %>%
  xtable(label = "tab:theta", 
         caption = "Parameters used to fit a test case with $V = H = 4$. This parameter vector was chosen as a sampled value of $\\boldsymbol \\theta$ that was not near the convex hull of the sufficient statistics for a grid point in Figure \\ref{fig:degen-plots} with $< 5$\\% near-degeneracy.", align = c("l", rep(c("r", "r"), times = 3))) %>%
  print(sanitize.text.function=function(x){x}, include.rownames = FALSE, comment = FALSE)

@@

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|}
\hline 
Method & Hyperparameter & Value \\ 
\hline \hline
\multirow{2}{*}{BwTPLV} & $C_1$ & $\frac{C}{n}\frac{1}{\nh + \nv}$ \\
 & $C_2$ & $\frac{C}{n}\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNLV} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNML} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\end{tabular}
\caption{The values used for the hyperparameters for all three fitting methods. A rule of thumb is imposed which decreases prior variances for the model parameters as the size of the model increases and also shrinks $\boldsymbol \theta_{interaction}$ more than $\boldsymbol \theta_{main}$. The common $C$ defining $C_1$ and $C_2$  in the BwTPLV method is chosen by tuning.}
\label{tab:hyperparam}
\end{table}

@@

```{r m_eff_table, results='asis'}
M_eff_table %>%
  gather(Image, M_eff, -Model) %>%
  spread(Model, M_eff) %>%
  mutate(Outcome = as.numeric(Image)) %>%
  select(Outcome, BwTNLV, BwTNML) %>%
  arrange(Outcome) %>%
  mutate(Outcome = as.character(Outcome)) -> M_eff_tmp

cbind(M_eff_tmp[1:8,], M_eff_tmp[9:16,]) %>%
  xtable(label = "tab:m-eff", 
         digits = 2, caption = paste0("The effective sample sizes for a chain of length $M = 1000$ regarding all $16$ probabilities for possible vector outcomes of visibles. BwTNLV would require at least $", round(min(M_eff_tmp[, "BwTNML"])/min(M_eff_tmp[, "BwTNLV"]), 1), "$ times as many MCMC iterations to achieve the same amount of effective information about the posterior distribution.")) %>%
  print(include.rownames = FALSE, comment = FALSE)
```

The effective sample sizes for a chain of length $M = 1000$ for inference about each of the $2^4 = 16$ model probabilities are presented in Table \ref{tab:m-eff}. These range from  $`r round(min(M_eff_tmp[, "BwTNML"]), 2)`$ to $`r round(max(M_eff_tmp[, "BwTNML"]), 2)`$ for BwTNML, while BwTNLV only yields between $`r round(min(M_eff_tmp[, "BwTNLV"]), 2)`$ and $`r round(max(M_eff_tmp[, "BwTNLV"]), 2)`$ effective draws. Thus, BwTNLV would require at least $`r round(min(M_eff_tmp[, "BwTNML"])/min(M_eff_tmp[, "BwTNLV"]), 1)`$ times as many iterations to be run of the MCMC chain in order to achieve the same amount of effective information about the posterior distribution. For this reason, consistent with the ACF results in Figure \ref{fig:acf}, BwTNLV does not seem to be an effective method for fitting the RBM if computing resources are at all limited.
